The Two-Column Specht Module of the Hecke Algebra of $S_n$

### 2019-07-04 tex/Crossingless.tex
Scaled im(1 + T_i) ~ W_{2n - 2} iso; added loop to corresponding figure.

\section{Irreducibility} 
We can now begin by proving that $\Res_{\SH_{k,p}(S_{2n - 1})}^{\SH_{k,p}(S_{2n})} W$ is irreducible;
in partuclar, this implies that $W$ itself is irreducible, as an $\SH$-subrepresentation is a $\SH_{k,p}(S_{2n - 1})$-subrepresentation.
\begin{proposition}
  Set $\SH' := \SH_{k,p}(S_{2n - 2}) \subset \SH$.
  Then, $\Res_{\SH'}^{\SH} W$ is irreducible if $q$ is generic or a primitive $e$'th root of unity for $e \geq n + 1$ or $e = 2$.
\end{proposition}
\begin{proof}
  We will prove the equivalent condition that each vector in $w \in W$ is \emph{cyclic}, i.e. $\SH' w = W$.

@@

  Since the image of $(1 + T_i)$ has arc $(i,i+1)$, it is isomorphic as a vector space to $W_{2n}$ with $2n - 2$ vertices.\footnote{This ismorphism ``ignores'' the arc $(i,i+1)$.}
  This isomorphism carries $w_1$ to an equivalent vector $w'_1$ in $W_{2n - 2}$ containing all length-2 arcs.
  Further, all actions of $(1 + T'_j) \in \SH_{k,p}(S_{2n - 4})$ act identically (through the isomorphism) with simple transpositions other than $1 + T'_i$, which acts equivalently to $(1 + T_i)(1 + T_{i + 1})(1 + T_{i - 1})$ as illustrated in Figure \ref{bigloop};
  hence, if the image of $(1 + T_i)w$ in $W_{2n - 2}$ generates an ideal containing $w'_1$, then $(1 + T_i)w$ generates an ideal containing $w_1$, and hence $w$ is cyclic.

  \begin{figure}[b]
  \[
    \Action{6}{1/6, 2/5, 3/4}{2}{1/6, 2/3, 4/5 }{q^{1/2}} \hspace{20pt} 
    \IsoAction{8}{1/8, 2/7, 3/4, 5/6}{3}{1/8, 2/5, 3/4, 6/7}{q^{1/2}}
  \]
  \caption{The correspondence between the action of $(1 + T_2)$ on $w'_5 \in W_6$ and the action of $(1 + T_3)(1 + T_4)(1 + T_2)$ on the corresponding vector in $W_8$ having arc $(3,4)$.}
  \label{bigloop}
  \end{figure}

>>>>

\newpage
\section{Irreducibility} 
We can now begin by proving that $\Res_{\SH_{k,p}(S_{2n - 1})}^{\SH_{k,p}(S_{2n})} W$ is irreducible;
in partuclar, this implies that $W$ itself is irreducible, as an $\SH$-subrepresentation is a $\SH_{k,p}(S_{2n - 1})$-subrepresentation.
\begin{proposition}
  Set $\SH' := \SH_{k,p}(S_{2n - 2}) \subset \SH$.
  Then, $\Res_{\SH'}^{\SH} W$ is irreducible if $q$ is generic or a primitive $e$'th root of unity for $e \geq n + 1$ or $e = 2$.
\end{proposition}
\begin{proof}
  We will prove the equivalent condition that each vector in $w \in W$ is \emph{cyclic}, i.e. $\SH' w = W$.

@@

  Since the image of $(1 + T_i)$ has arc $(i,i+1)$, it is isomorphic as a vector space to $W_{2n}$ with $2n - 2$ vertices.\footnote{This ismorphism ``ignores'' the arc $(i,i+1)$.}
  This isomorphism carries $w_1$ to an equivalent vector $w'_1$ in $W_{2n - 2}$ containing all length-2 arcs.
  Further, all actions of $(1 + T'_j) \in \SH_{k,p}(S_{2n - 4})$ act identically (through the isomorphism) with simple transpositions other than $1 + T'_i$, which acts equivalently to $q^{-1}(1 + T_i)(1 + T_{i + 1})(1 + T_{i - 1})$ as illustrated in Figure \ref{bigloop};
  hence, if the image of $(1 + T_i)w$ in $W_{2n - 2}$ generates an ideal containing $w'_1$, then $(1 + T_i)w$ generates an ideal containing $w_1$, and hence $w$ is cyclic.

  \begin{figure}[b]
  \[
    \Action{6}{1/6, 2/5, 3/4}{2}{1/6, 2/3, 4/5 }{q^{1/2}} \hspace{20pt} 
    \IsoAction{8}{1/8, 2/7, 3/4, 5/6}{3}{1/8, 2/5, 3/4, 6/7}{q^{1/2}}
  \]
  \[
    \Action{4}{1/4, 2/3}{2}{1/4,2/3}{(1 + q)} \hspace{20pt}
    \IsoAction{6}{1/6, 2/5, 3/4}{3}{1/6, 2/5, 3/4}{(1 + q)}
  \]
  \caption{The correspondence between the action of $(1 + T_2)$ on $w'_5 \in W_6$ and the action of $q^{-1}(1 + T_3)(1 + T_4)(1 + T_2)$ on the corresponding vector in $W_8$ having arc $(3,4)$ first, then on $w'_2 \in W_4$.
  This demonstrates that the action works with and without creating a loop.
  }
  \label{bigloop}
  \end{figure}
